---
title: "SNP FILTERING"
output:
  html_document:
    toc: yes
  html_notebook:
    code_folding: hide
    df_print: paged
    highlight: kate
    theme: yeti
    toc: yes
---

```{r load libraries, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

source("scr/libraries.R")
source("scr/ggplot.R")
source("scr/VCFfilterstats.R")
source("scr/xtrafunctions.R")

```

# Bonnethead sharks

## Raw data stats

### Query stats

Query raw stats using vcftools.

```{bash}

# depth indv/locus
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --depth
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --site-mean-depth
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --missing-indv
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --indv-freq-burden
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --freq2
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --singletons
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --012

# heterozygosity per individual
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --het

# SNP call quality
vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out results/BONNET_raw --site-quality

```

### Visualize

```{r stats raw, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_raw <- read.ind.stats(dir = "results", vcf = "BONNET_raw")

loc_stats_raw <- read.loc.stats(dir = "results", vcf = "BONNET_raw")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_raw, aes(x = MISS_BONNET_raw)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MISS_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_raw, aes(x = Fis_BONNET_raw)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(Fis_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_raw, aes(x = MEAN_DEPTH_BONNET_raw)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_raw, aes(x = MEAN_DEPTH_BONNET_raw, y = MISS_BONNET_raw)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(MISS_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_raw, aes(x = Fis_BONNET_raw, y = MISS_BONNET_raw)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(Fis_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(MISS_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_raw, aes(x = Fis_BONNET_raw, y = MEAN_DEPTH_BONNET_raw)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(Fis_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(MEAN_DEPTH_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_raw, aes(x = MISS_BONNET_raw)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MISS_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_raw, aes(x = MEAN_DEPTH_BONNET_raw)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_raw, aes(x = MEAN_DEPTH_BONNET_raw, y = MISS_BONNET_raw)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(MISS_BONNET_raw, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_raw %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_raw %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_raw) %>%
  ggplot() +
  geom_point(aes(x = n, y = MEAN_DEPTH_BONNET_raw)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_raw.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_raw$MEAN_DEPTH_BONNET_raw, site_qual$QUAL) %>%
  rename(depth = loc_stats_raw.MEAN_DEPTH_BONNET_raw, qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m1 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_raw)` individuals and `r nrow(loc_stats_raw)` loci.


## Filtering scheme 1

### Step 1: Remove loci with genotype call rate < 95%:

```{bash}

vcftools --vcf data/BONNET/TotalRawSNPs.vcf --out data/BONNET//BONNET_geno95 --max-missing 0.95 --recode --recode-INFO-all

```

Decompose variants and retain only SNPs.

```{bash}

vcfallelicprimitives data/BONNET/BONNET_geno95.recode.vcf --keep-info --keep-geno > data/BONNET/BONNET_SNPs.vcf
vcftools --vcf data/BONNET/BONNET_SNPs.vcf --out data/BONNET/BONNET_geno95SNPs --remove-indels --recode --recode-INFO-all


vcftools --vcf data/BONNET/BONNET_geno95SNPs.recode.vcf --out results/BONNET_geno95SNPs --missing-indv

```

### Step 2: Identify individuals with > 25% missing data:

```{r}

imissing <- read.table("results/BONNET_geno95SNPs.imiss",
                       header = TRUE, stringsAsFactors = FALSE)

LQ_indv <- imissing %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

ggplot(imissing, aes(x = F_MISS)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(F_MISS, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.25),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

write.table(LQ_indv, "data/BONNET//LQ_Ind_FIL-1", 
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals:

```{bash}

vcftools --vcf data/BONNET/BONNET_geno95SNPs.recode.vcf --out data/BONNET//BONNET_FIL-1 --remove data/BONNET//LQ_Ind_FIL-1 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --depth
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --missing-indv
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --freq2
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --singletons
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_FIL-1.recode.vcf --out results/BONNET_FIL-1 --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 1, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_1 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-1")

loc_stats_FIL_1 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-1")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_1, aes(x = `MISS_BONNET_FIL-1`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_1, aes(x = `Fis_BONNET_FIL-1`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_1, aes(x = `MEAN_DEPTH_BONNET_FIL-1`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_1, aes(x = `MEAN_DEPTH_BONNET_FIL-1`, y = `MISS_BONNET_FIL-1`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_1, aes(x = `Fis_BONNET_FIL-1`, y = `MISS_BONNET_FIL-1`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_1, aes(x = `Fis_BONNET_FIL-1`, y = `MEAN_DEPTH_BONNET_FIL-1`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_1, aes(x = `MISS_BONNET_FIL-1`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_1, aes(x = `MEAN_DEPTH_BONNET_FIL-1`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_1, aes(x = `MEAN_DEPTH_BONNET_FIL-1`, y = `MISS_BONNET_FIL-1`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-1`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_1 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_1 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_1) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-1`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-1.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_1$`MEAN_DEPTH_BONNET_FIL-1`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_1..MEAN_DEPTH_BONNET_FIL.1., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m2 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_1)` individuals and `r nrow(loc_stats_FIL_1)` loci.


## Filtering scheme 2

### Step 1: Identify individuals with > 25% missing data:

```{r, fig.height=4, fig.width=5}

imissing <- read.table("results/BONNET_raw.imiss",
                       header = TRUE, stringsAsFactors = FALSE)

ggplot(imissing, aes(x = F_MISS)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(F_MISS, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.25),
                 color = "darkblue", linetype = "dashed", size = 1) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x = "missing data per indv") +
  theme_standard

LQ_indv <- imissing %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_FIL-2", 
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals:

```{bash}

vcftools --vcf data/BONNET//TotalRawSNPs.vcf --out data/BONNET//BONNET_miss25 --remove data/BONNET//LQ_Ind_FIL-2 --recode --recode-INFO-all

```

### Step 2: Remove loci with genotype call rate < 95%, decompose variants and retain only SNPs.

```{bash}

vcftools --vcf data/BONNET/BONNET_miss25.recode.vcf --out data/BONNET//BONNET_miss25geno95 --max-missing 0.95 --recode --recode-INFO-all

vcfallelicprimitives data/BONNET//BONNET_miss25geno95.recode.vcf --keep-info --keep-geno > data/BONNET//BONNET_miss25geno95SNPs.vcf

vcftools --vcf data/BONNET/BONNET_miss25geno95SNPs.vcf --out data/BONNET//BONNET_FIL-2 --remove-indels --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --depth
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --missing-indv
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --freq2
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --singletons
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_FIL-2.recode.vcf --out results/BONNET_FIL-2 --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 2, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_2 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-2")

loc_stats_FIL_2 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-2")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_2, aes(x = `MISS_BONNET_FIL-2`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_2, aes(x = `Fis_BONNET_FIL-2`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_2, aes(x = `MEAN_DEPTH_BONNET_FIL-2`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_2, aes(x = `MEAN_DEPTH_BONNET_FIL-2`, y = `MISS_BONNET_FIL-2`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_2, aes(x = `Fis_BONNET_FIL-2`, y = `MISS_BONNET_FIL-2`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_2, aes(x = `Fis_BONNET_FIL-2`, y = `MEAN_DEPTH_BONNET_FIL-2`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_2, aes(x = `MISS_BONNET_FIL-2`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_2, aes(x = `MEAN_DEPTH_BONNET_FIL-2`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_2, aes(x = `MEAN_DEPTH_BONNET_FIL-2`, y = `MISS_BONNET_FIL-2`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-2`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_2 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_2 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_2) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-2`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-2.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_2$`MEAN_DEPTH_BONNET_FIL-2`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_2..MEAN_DEPTH_BONNET_FIL.2., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m3 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_2)` individuals and `r nrow(loc_stats_FIL_2)` loci.


## Filtering scheme 3

### Step 1: Filter loci

Filter loci with minimum SNP call quality per locus < 20, minimum genotype depth < 5, then remove loci with mean minimum depth < 15, filter loci with minor allele count < 3, finally apply genotype call rate > 95%.

```{bash}

##
vcftools --vcf data/BONNET/TotalRawSNPs.vcf --out data/BONNET/BONNET_minQ20minDP5 --minQ 20 --minDP 5 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --missing-indv
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --freq2
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --singletons
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out results/BONNET_minQ20minDP5 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf--out results/BONNET_minQ20minDP5 --site-quality


##
vcftools --vcf data/BONNET/BONNET_minQ20minDP5.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10 --min-meanDP 15 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --missing-indv
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --freq2
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --singletons
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out results/BONNET_minQ20minDP5meanDP10 --site-quality


##
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10.recode.vcf --out data/BONNET/BONNET_minQ20minDP5meanDP10mac3 --mac 3 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --missing-indv
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --freq2
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --singletons
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3 --site-quality


##
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95 --max-missing 0.95 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --missing-indv
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --freq2
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --singletons
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out results/BONNET_minQ20minDP5meanDP10mac3geno95 --site-quality

```

Decompose variants and retain only SNPs.

```{bash}

vcfallelicprimitives data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --keep-info --keep-geno > data/BONNET/SNPs.vcf

vcftools --vcf data/BONNET/SNPs.vcf --out data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.SNPs --remove-indels --recode --recode-INFO-all

vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.SNPs --out results/BONNET_minQ20minDP5meanDP10mac3geno95.SNPs  --missing-indv

```

### Step 2: Identify individuals with > 25% missing data:

```{r}

imissing <- read.table("results/BONNET_minQ20minDP5meanDP10mac3geno95.imiss",
                       header = TRUE, stringsAsFactors = FALSE)

ggplot(imissing, aes(x = F_MISS)) +
  geom_histogram(binwidth = .05, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(F_MISS, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.25),
                 color = "darkblue", linetype = "dashed", size = 1) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x = "missing data per indv") +
  theme_standard

LQ_indv <- imissing %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_FIL-3", 
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals:

```{bash}

vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10mac3geno95.recode.vcf --out data/BONNET/BONNET_FIL-3 --remove data/BONNET/LQ_Ind_FIL-3 --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --depth
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --site-mean-depth
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --missing-indv
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --freq2
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --singletons
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_FIL-3.recode.vcf --out results/BONNET_FIL-3 --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 3, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_3 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-3")

loc_stats_FIL_3 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-3")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_3, aes(x = `MISS_BONNET_FIL-3`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_3, aes(x = `Fis_BONNET_FIL-3`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_3, aes(x = `MEAN_DEPTH_BONNET_FIL-3`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_3, aes(x = `MEAN_DEPTH_BONNET_FIL-3`, y = `MISS_BONNET_FIL-3`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_3, aes(x = `Fis_BONNET_FIL-3`, y = `MISS_BONNET_FIL-3`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_3, aes(x = `Fis_BONNET_FIL-3`, y = `MEAN_DEPTH_BONNET_FIL-3`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_3, aes(x = `MISS_BONNET_FIL-3`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_3, aes(x = `MEAN_DEPTH_BONNET_FIL-3`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_3, aes(x = `MEAN_DEPTH_BONNET_FIL-3`, y = `MISS_BONNET_FIL-3`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-3`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_3 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_3 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_3) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-3`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-3.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_3$`MEAN_DEPTH_BONNET_FIL-3`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_3..MEAN_DEPTH_BONNET_FIL.3., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m4 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_3)` individuals and `r nrow(loc_stats_FIL_3)` loci.

## Filtering scheme 4

### Step 1: Filter LQ loci

Remove SNPs with quality score < 20, minor allele count < 3, minimum depth per genotype < 5, minimum mean depth < 15, mac < 3. 

Then remove loci with genotype call rate < 50%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno50 --max-missing 0.5 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno50.recode.vcf --out results/BONNET_it1 --missing-indv

```

### Step 2: Missing data loci/indv

Identify individuals with > 90% missing data

```{r}

imiss <- read.table("results/BONNET_it1.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .9, color = "darkblue", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.9) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it1a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno50.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno50ind90 --remove data/BONNET//LQ_Ind_it1a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 60%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno50ind90.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno60ind90 --max-missing 0.6 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno60ind90.recode.vcf --out results/BONNET_it2 --missing-indv

```

Identify individuals with > 70% missing data

```{r}

imiss <- read.table("results/BONNET_it2.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .7, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.7) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it2a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10mac3geno60ind90.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno60ind70 --remove data/BONNET//LQ_Ind_it2a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 70%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno60ind70.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind70 --max-missing 0.7 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind70.recode.vcf --out results/BONNET_it3 --missing-indv

```

Identify individuals with > 50% missing data

```{r}

imiss <- read.table("results/BONNET_it3.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .5, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.5) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it3a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind70.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50 --remove data/BONNET//LQ_Ind_it3a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 80%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno80ind50 --max-missing 0.8 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno80ind50.recode.vcf --out results/BONNET_it4 --missing-indv

```

Identify individuals with > 30% missing data

```{r}

imiss <- read.table("results/BONNET_it4.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .3, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.3) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it4a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno80ind50.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno80ind30 --remove data/BONNET//LQ_Ind_it4a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 90%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno80ind30.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno90ind30 --max-missing 0.9 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno90ind30.recode.vcf --out results/BONNET_it5 --missing-indv

```

Identify individuals with > 25% missing data

```{r}

imiss <- read.table("results/BONNET_it5.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .25, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

write.table(LQ_indv, "results/LQ_Ind_it5a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno90ind30.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno90ind25 --remove results/LQ_Ind_it5a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 95%; decompose variants and retain only SNPs

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno90ind25.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno95ind25 --max-missing 0.95 --recode --recode-INFO-all

vcfallelicprimitives data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno95ind25.recode.vcf --keep-info --keep-geno > data/BONNET//SNPs.vcf

vcftools --vcf data/BONNET//SNPs.vcf --out data/BONNET//BONNET_FIL-4 --remove-indels --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --depth
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --site-mean-depth
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --missing-indv
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --indv-freq-burden
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --freq2
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --singletons
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4 --012

# heterozygosity per individual
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4  --het

# SNP call quality
vcftools --vcf data/BONNET//BONNET_FIL-4.recode.vcf --out results/BONNET_FIL-4  --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 4, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_4 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-4")

loc_stats_FIL_4 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-4")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_4, aes(x = `MISS_BONNET_FIL-4`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_4, aes(x = `Fis_BONNET_FIL-4`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_4, aes(x = `MEAN_DEPTH_BONNET_FIL-4`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_4, aes(x = `MEAN_DEPTH_BONNET_FIL-4`, y = `MISS_BONNET_FIL-4`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_4, aes(x = `Fis_BONNET_FIL-4`, y = `MISS_BONNET_FIL-4`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_4, aes(x = `Fis_BONNET_FIL-4`, y = `MEAN_DEPTH_BONNET_FIL-4`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_4, aes(x = `MISS_BONNET_FIL-4`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_4, aes(x = `MEAN_DEPTH_BONNET_FIL-4`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_4, aes(x = `MEAN_DEPTH_BONNET_FIL-4`, y = `MISS_BONNET_FIL-4`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-4`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_4 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_4 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_4) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-4`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-4.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_4$`MEAN_DEPTH_BONNET_FIL-4`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_4..MEAN_DEPTH_BONNET_FIL.4., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m5 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_4)` individuals and `r nrow(loc_stats_FIL_4)` loci.

## Filtering scheme 5

### Step 1: Filter LQ loci

Remove SNPs with quality score < 20, minor allele count < 3, minimum depth per genotype < 5, minimum mean depth < 15, mac < 3. 

### Step 2: Missing data indvidiuals/loci 

Retain loci with genotype call rate > 70% and individuals with missing data < 50%.

### Step 3: Info flag filtering

#### Allele balance

AB: Allele balance at heterozygous sites: a number between 0 and 1 representing the ratio of reads showing the reference allele to all reads, considering only reads from individuals called as heterozygous

```{bash query INFO stats I, eval=FALSE, include=FALSE}

cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind70.recode.vcf | grep -oe "AB=[[:digit:]].[[:digit:]][[:digit:]][[:digit:]]" | sed -s 's/AB=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.AB

# allele balance
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "AB=[[:digit:]].[[:digit:]][[:digit:]][[:digit:]]" | sed -s 's/AB=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.AB

# site Depth
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "DP=[0-9]*" | sed -s 's/DP=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.DEPTH

# quality score
mawk '!/#/' BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | cut -f1,2,6 > BONNET_minQ20mac3minDP5meanDP10geno70ind50.QUAL

# mapping quality
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "MQM=[0-9]*" | sed -s 's/MQM=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQM
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "MQMR=[0-9]*" | sed -s 's/MQMR=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQMR

# strand balance
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SAF=[0-9]*" | sed -s 's/SAF=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.SAF
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SAR=[0-9]*" | sed -s 's/SAR=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.SAR
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SRF=[0-9]*" | sed -s 's/SRF=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.SRF
cut -f8 BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SRR=[0-9]*" | sed -s 's/SRR=//g' > BONNET_minQ20mac3minDP5meanDP10geno70ind50.SRR

```

Allele balance is the ratio of reads for reference allele to all reads, considering only reads from individuals called as heterozygous. Values range from 0 - 1; allele balance (for real loci) should be approx. 0.5. Filter contigs SNPs for which the with allele balance < 0.25 and > 0.75.

```{r plot AB}

read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.AB",
           col.names = "AB", stringsAsFactors = FALSE) %>%
  ggplot(aes(x = AB)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(xintercept = 0.5, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.2, color = "darkblue", linetype = "dashed") +
  geom_vline(xintercept = 0.8, color = "darkblue", linetype = "dashed") +
  labs(x = "Allele balance (red BONNET)") +
  theme_standard

```

Filter contigs with SNP calls with AB > 0.2, AB > 0.8; retain loci very close to 0 (retain loci that are fixed variants). Remove genotypes if the quality sum of the reference or alternate allele was 0.

```{bash, eval=FALSE, include=FALSE}

# this works... not sure why cant write to file in the folder
vcffilter -s -f "AB > 0.2 & AB < 0.8 | AB < 0.01 | AB > 0.99" -s -g "QR > 0 | QA > 0"  BONNET_minQ20mac3minDP5meanDP10geno70ind50.recode.vcf >  BONNET_minQ20mac3minDP5meanDP10geno70ind50AB.recode.vcf 

mawk '!/#/' BONNET_minQ20mac3minDP5meanDP10geno70ind50AB.recode.vcf | wc -l

```

Remaining SNPs: 29.801.

#### Quality/depth ratio

Compare quality/depth.

```{r, fig.height=10, fig.width=8, message=FALSE, warning=FALSE}

# depth
depth <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.DEPTH",
                    col.names = "depth")

# quality score
qual <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.QUAL",
                   col.names = c("locus", "pos", "qual"))

df <- bind_cols(qual, depth) %>%
  mutate(ratio = qual/depth)

ggplot(df, aes(x = ratio)) + 
  geom_histogram(binwidth = 0.1, color = "black", fill = "grey95") +
  geom_vline(xintercept = 0.2, color = "darkred", linetype = "dashed") +
  geom_vline(xintercept = 0.5, color = "darkred", linetype = "dashed") +
  labs(x = "ratio depth/quality (red BONNET)", y = "no. of loci") +
  theme_standard

```

Remove loci with quality/depth ratio < 0.2

```{bash}

vcffilter -s -f "QUAL / DP > 0.2" BONNET_minQ20mac3minDP5meanDP10geno70ind50AB.recode.vcf > BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTH.recode.vcf

mawk '!/#/' BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTH.recode.vcf | wc -l

```

Remaining SNPs: 28,570.

#### ratio mapping quality

Remove loci based on ratio of mapping quality for reference and alternate allele, i.e. sites that have a high discrepancy between the mapping qualities of two alleles.

```{r plot map qual ratios}

temp <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQM", col.names = "MQM")

mapqual <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQMR", col.names = "MQMR")

mapqual <- bind_cols(mapqual, temp) %>%
  mutate(ratio = MQM/MQMR)

filter <- mapqual %>%
  filter(ratio < 0.25 | ratio > 1.75)

ggplot(mapqual, aes(x = MQM, y = MQMR)) +
  geom_point(shape = 1) + 
  geom_abline(intercept = 0, slope = 1, size = 1, color = "red", linetype = "dashed") +
  geom_abline(intercept = 0, slope = 4, size = 1, color = "darkblue", linetype = "dashed") +
  geom_abline(intercept = 0, slope = 0.571, size = 1, color = "darkblue", linetype = "dashed") +
  geom_point(data = filter, aes(x = MQM, y = MQMR), shape = 1, color = "red") +
  scale_x_continuous(limits = c(0, 65)) + 
  scale_y_continuous(limits = c(0, 65)) +
  labs(x = "mean mapping quality alt allele (red BONNET)", y = "mean mapping quality ref allele ") +
  theme_standard

```

Filter loci with mapping quality ratio < 0.25 and > 1.75.

```{bash filter map qual ratio}

vcffilter -s -f "MQM / MQMR > 0.25 & MQM / MQMR < 1.75" BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTH.recode.vcf > BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf

mawk '!/#/' BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf | wc -l

```

Remaining SNPs: 28.132.

#### Maximum depth & Quality

Identify distribution of depth (based on original data set) to identify loci with excess coverage.

Original number of individuals in data set is `r nrow(ind_stats_raw)` (INFO flags in filtered data set are are based on original number of individuals in data set).

Calculate average depth and standard deviation:

```{r plot depth vs qual}

# depth
depth <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.DEPTH",
                    col.names = "depth")

# quality score
qual <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.QUAL",
                   col.names = c("locus", "pos", "qual"))

# mean depth
mean_depth <- mean(depth$depth)

# standard deviation
std <- sd(depth$depth)

# mode
mode <- Mode(depth$depth)

cutoff <- sum(mean_depth + (2*std))

# identify SNP sites with depth > mean depth + 1 standard deviation and quality score < 2x the depth at that site
temp <- bind_cols(qual, depth) %>%
  filter(depth > cutoff) %>%
  filter(qual < 2*depth) %>%
  select(locus)

write.table(temp, "data/BONNET//DEPTH.lowQloci", col.names = FALSE, row.names = FALSE, quote = FALSE)

df <- bind_cols(qual, depth) %>%
  mutate(qualcutoff = 2*depth)

removeloc <- df %>%
  filter(depth > cutoff) %>%
  filter(qual < 2*depth)

ggplot(df, aes(x = depth, y = qual)) +
  geom_point(shape = 1) +
  geom_point(data = removeloc, aes(x = depth, y = qual), shape = 1, color = "red") +
  geom_line(data = df, aes(x = depth, y = qualcutoff), color = "blue",  linetype = "dashed", size = 1) +
  geom_vline(xintercept = cutoff, color = "blue", linetype = "dashed", size = 1) +
  labs(x = "red BONNET") +
  theme_standard

```

Mean depth per locus (across all indivuals) is `r mean_depth` and the standard deviation is `r std`. 

Filter SNP site with depth > mean depth + 1 standard deviation = `r sum(mean_depth + 2*std)` and that have quality scores < 2x the depth at that site and output depth per site.

```{bash filter depth qual, eval=FALSE, include=FALSE}

vcftools --vcf  data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf --exclude-positions data/BONNET//DEPTH.lowQloci --recode --recode-INFO-all --out data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQMQUAL

vcftools --vcf data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.recode.vcf --out data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQMQUAL --site-mean-depth

```

Compare the distribution of mean depth per site averaged across individuals to determine cut-off value of sites with excessively high depth indicative of paralogs/multicopy loci.

```{r plot depth dist, message=FALSE, warning=FALSE}

# calculate mean depth per site (177 individuals)
depth <- read.table("data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.ldepth.mean",
           header = TRUE, stringsAsFactors = FALSE)
  
# mean mean depth
mean <- mean(depth$MEAN_DEPTH)

# standard deviation
std <- sd(depth$MEAN_DEPTH)

# mode
mode <- Mode(depth$MEAN_DEPTH)

cutoff <- sum(mean + (2*std))

ggplot(depth, aes(x = MEAN_DEPTH)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mode),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cutoff),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per site") +
  theme_standard

```

Choose cut-off for maximum mean read depth = 120 (calculate mean + 2x std: `r sum(mean + (2*std))`)

```{bash filter max depth}

vcftools --vcf  data/BONNET/BONNET_minQ20mac3minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.recode.vcf --max-meanDP 120 --exclude-positions data/BONNET//DEPTH.lowQloci --recode --recode-INFO-all --out data/BONNET/BONNET_dDocent

# depth indv/locus
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --depth
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --site-mean-depth
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --missing-indv
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --indv-freq-burden
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --freq2
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --singletons
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --012

# heterozygosity per individual
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --het

# SNP call quality
vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out results/BONNET_dDocent --site-quality

```

#### Plot stats after dDocent filters

```{r stats dDocent, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_dDocent <- read.ind.stats(dir = "results", vcf = "BONNET_dDocent")

loc_stats_dDocent <- read.loc.stats(dir = "results", vcf = "BONNET_dDocent")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_dDocent, aes(x = `MISS_BONNET_dDocent`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocent`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocent`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocent`, y = `MISS_BONNET_dDocent`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocent`, y = `MISS_BONNET_dDocent`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocent`, y = `MEAN_DEPTH_BONNET_dDocent`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_dDocent, aes(x = `MISS_BONNET_dDocent`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocent`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocent`, y = `MISS_BONNET_dDocent`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocent`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_dDocent %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_dDocent %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_dDocent) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_dDocent`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_dDocent.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_dDocent$`MEAN_DEPTH_BONNET_dDocent`, site_qual$QUAL) %>%
  rename(depth = loc_stats_dDocent.MEAN_DEPTH_BONNET_dDocent, qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m6 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

### Step 4: Filter Missing data loci/indv

Remove loci with genotype call rate < 80%.

```{bash}

vcftools --vcf data/BONNET//BONNET_dDocent.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind50 --max-missing 0.8 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind50.recode.vcf --out results/BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind50 --missing-indv

```

Identify individuals with > 30% missing data

```{r}

imiss <- read.table("results/BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind50.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .3, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.3) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_FIL-5a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind50.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind30 --remove data/BONNET//LQ_Ind_FIL-5a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 90%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno80ind30.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind30 --max-missing 0.9 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind30.recode.vcf --out results/BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind30 --missing-indv

```

Identify individuals with > 25% missing data

```{r}

imiss <- read.table("results/BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind30.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .25, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

write.table(LQ_indv, "results/LQ_Ind_FIL-5b",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind30.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind25 --remove results/LQ_Ind_FIL-5b --recode --recode-INFO-all

```

Remove loci with genotype call rate < 95%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno90ind25.recode.vcf --out data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno95ind25 --max-missing 0.95 --recode --recode-INFO-all

vcfallelicprimitives data/BONNET//BONNET_minQ20mac3minDP5meanDP10dDocentgeno95ind25.recode.vcf --keep-info --keep-geno > data/BONNET//SNPs.vcf

vcftools --vcf data/BONNET//SNPs.vcf --out data/BONNET//BONNET_FIL-5 --remove-indels --recode --recode-INFO-all

# depth indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --depth
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --site-mean-depth
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --missing-indv
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --indv-freq-burden
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --freq2
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --singletons
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --012

# heterozygosity per individual
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --het

# SNP call quality
vcftools --vcf data/BONNET//BONNET_FIL-5.recode.vcf --out results/BONNET_FIL-5 --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 5, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_5 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-5")

loc_stats_FIL_5 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-5")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_5, aes(x = `MISS_BONNET_FIL-5`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_5, aes(x = `Fis_BONNET_FIL-5`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_5, aes(x = `MEAN_DEPTH_BONNET_FIL-5`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_5, aes(x = `MEAN_DEPTH_BONNET_FIL-5`, y = `MISS_BONNET_FIL-5`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_5, aes(x = `Fis_BONNET_FIL-5`, y = `MISS_BONNET_FIL-5`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_5, aes(x = `Fis_BONNET_FIL-5`, y = `MEAN_DEPTH_BONNET_FIL-5`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_5, aes(x = `MISS_BONNET_FIL-5`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_5, aes(x = `MEAN_DEPTH_BONNET_FIL-5`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_5, aes(x = `MEAN_DEPTH_BONNET_FIL-5`, y = `MISS_BONNET_FIL-5`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-5`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_5 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_5 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_5) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-5`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-5.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_5$`MEAN_DEPTH_BONNET_FIL-5`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_5..MEAN_DEPTH_BONNET_FIL.5., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m7 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_5)` individuals and `r nrow(loc_stats_FIL_5)` loci.

## Filtering scheme 6

### Step 1: Filter LQ loci

Remove SNPs with quality score < 20, minimum depth per genotype < 5, and minimum mean depth < 15. 

Then remove loci with genotype call rate < 50%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10.recode.vcf --out data/BONNET//BONNET_minQminDP5meanDP10geno50 --max-missing 0.5 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQminDP5meanDP10geno50.recode.vcf --out results/BONNET_it1b --missing-indv

```

### Step 2: Missing data indvidiuals/loci 

Retain loci with genotype call rate > 70% and individuals with missing data < 50%.

#### Identify individuals with > 90% missing data

```{r}

imiss <- read.table("results/BONNET_it1b.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .9, color = "darkblue", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.9) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it1b",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQminDP5meanDP10geno50.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10geno50ind90 --remove data/BONNET//LQ_Ind_it1b --recode --recode-INFO-all

```

Remove loci with genotype call rate < 60%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno50ind90.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10geno60ind90 --max-missing 0.6 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno60ind90.recode.vcf --out results/BONNET_it2b --missing-indv

```

Identify individuals with > 70% missing data

```{r}

imiss <- read.table("results/BONNET_it2b.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .7, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.7) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it2b",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno60ind90.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10geno60ind70 --remove data/BONNET//LQ_Ind_it2b --recode --recode-INFO-all

```

Remove loci with genotype call rate < 70%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno60ind70.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind70 --max-missing 0.7 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind70.recode.vcf --out results/BONNET_it3b --missing-indv

```

Identify individuals with > 50% missing data

```{r}

imiss <- read.table("results/BONNET_it3b.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .5, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.5) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_it3b",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind70.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50 --remove data/BONNET//LQ_Ind_it3b --recode --recode-INFO-all

```

### Step 3: Info flag filtering

#### Query stats

```{bash query INFO stats II, eval=FALSE, include=FALSE}

# allele balance
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "AB=[[:digit:]].[[:digit:]][[:digit:]][[:digit:]]" | sed -s 's/AB=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.AB

# site Depth
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "DP=[0-9]*" | sed -s 's/DP=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.DEPTH

# quality score
mawk '!/#/' BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | cut -f1,2,6 > BONNET_minQ20minDP5meanDP10geno70ind50.QUAL

# mapping quality
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "MQM=[0-9]*" | sed -s 's/MQM=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.MQM
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "MQMR=[0-9]*" | sed -s 's/MQMR=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.MQMR

# strand balance
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SAF=[0-9]*" | sed -s 's/SAF=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.SAF
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SAR=[0-9]*" | sed -s 's/SAR=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.SAR
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SRF=[0-9]*" | sed -s 's/SRF=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.SRF
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "SRR=[0-9]*" | sed -s 's/SRR=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.SRR

# properly paired status
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "PAIRED=[0-9]\.[0-9]*" | sed -s 's/PAIRED=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.PAIRED
cut -f8 BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf | grep -oe "PAIREDR=[0-9]\.[0-9]*" | sed -s 's/PAIREDR=//g' > BONNET_minQ20minDP5meanDP10geno70ind50.PAIREDR

```

#### Allele balance

AB: Allele balance at heterozygous sites: a number between 0 and 1 representing the ratio of reads showing the reference allele to all reads, considering only reads from individuals called as heterozygous

Allele balance is the ratio of reads for reference allele to all reads, considering only reads from individuals called as heterozygous. Values range from 0 - 1; allele balance (for real loci) should be approx. 0.5. Filter contigs SNPs for which the with allele balance < 0.25 and > 0.75.

```{r plot AB II, fig.height=3, fig.width=4}

read.table("data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50.AB",
           col.names = "AB", stringsAsFactors = FALSE) %>%
  ggplot(aes(x = AB)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(xintercept = 0.5, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.2, color = "darkblue", linetype = "dashed") +
  geom_vline(xintercept = 0.8, color = "darkblue", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x = "Allele balance (BONNET)") +
  theme_standard

```

Filter contigs with SNP calls with AB > 0.2, AB > 0.8; retain loci very close to 0 (retain loci that are fixed variants). Remove genotypes if the quality sum of the reference or alternate allele was 0.

```{bash, eval=FALSE, include=FALSE}

# this works... not sure why cant write to file in the folder
vcffilter -s -f "AB > 0.2 & AB < 0.8 | AB < 0.01 | AB > 0.99" -s -g "QR > 0 | QA > 0"  BONNET_minQ20minDP5meanDP10geno70ind50.recode.vcf >  BONNET_minQ20minDP5meanDP10geno70ind50AB.recode.vcf 

mawk '!/#/' BONNET_minQ20minDP5meanDP10geno70ind50AB.recode.vcf | wc -l

```

Remaining SNPs: 50,938.

#### Quality/depth ratio

Compare quality/depth.

```{r, fig.height=10, fig.width=8, message=FALSE, warning=FALSE}

# depth
depth <- read.table("data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50.DEPTH",
                    col.names = "depth")

# quality score
qual <- read.table("data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50.QUAL",
                   col.names = c("locus", "pos", "qual"))

temp <- bind_cols(qual, depth) %>%
  mutate(ratio = qual/depth)

ggplot(temp, aes(x = ratio)) + 
  geom_histogram(binwidth = 0.1, color = "black", fill = "grey95") +
  geom_vline(xintercept = 0.2, color = "darkred", linetype = "dashed") +
  geom_vline(xintercept = 0.5, color = "darkred", linetype = "dashed") +
  labs(x = "ratio quality/depth (red BONNET)", y = "no. of loci") +
  theme_standard

```

Remove loci with quality/depth ratio < 0.2

```{bash}

vcffilter -s -f "QUAL / DP > 0.2" BONNET_minQ20minDP5meanDP10geno70ind50AB.recode.vcf > BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTH.recode.vcf

mawk '!/#/' BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTH.recode.vcf | wc -l

```

Remaining SNPs: 41,085.

#### ratio mapping quality

Remove loci based on ratio of mapping quality for reference and alternate allele, i.e. sites that have a high discrepancy between the mapping qualities of two alleles.

```{r plot map qual ratios II}

temp <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQM", col.names = "MQM")

mapqual <- read.table("data/BONNET//BONNET_minQ20mac3minDP5meanDP10geno70ind50.MQMR", col.names = "MQMR")

mapqual <- bind_cols(mapqual, temp) %>%
  mutate(ratio = MQM/MQMR)

filter <- mapqual %>%
  filter(ratio < 0.25 | ratio > 1.75)

ggplot(mapqual, aes(x = MQM, y = MQMR)) +
  geom_point(shape = 1) + 
  geom_abline(intercept = 0, slope = 1, size = 1, color = "red", linetype = "dashed") +
  geom_abline(intercept = 0, slope = 4, size = 1, color = "darkblue", linetype = "dashed") +
  geom_abline(intercept = 0, slope = 0.571, size = 1, color = "darkblue", linetype = "dashed") +
  geom_point(data = filter, aes(x = MQM, y = MQMR), shape = 1, color = "red") +
  scale_x_continuous(limits = c(0, 65)) + 
  scale_y_continuous(limits = c(0, 65)) +
  labs(x = "mean mapping quality alt allele (red BONNET)", y = "mean mapping quality ref allele ") +
  theme_standard

```

Filter loci with mapping quality ratio < 0.25 and > 1.75.

```{bash filter map qual ratio II}

vcffilter -s -f "MQM / MQMR > 0.25 & MQM / MQMR < 1.75" BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTH.recode.vcf > BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf

mawk '!/#/' BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf | wc -l

```

Remaining SNPs: 40,482.

#### Maximum depth & Quality

Identify distribution of depth (based on original data set) to identify loci with excess coverage.

Original number of individuals in data set is `r nrow(ind_stats_raw)` (INFO flags in filtered data set are are based on original number of individuals in data set).

Create file with the original site depth and quality score for each site:

Calculate average depth and standard deviation:

```{r plot depth vs qual II}

# depth
depth <- read.table("data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50.DEPTH",
                    col.names = "depth")

# quality score
qual <- read.table("data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50.QUAL",
                   col.names = c("locus", "pos", "qual"))

# mean depth
mean_depth <- mean(depth$depth)

# standard deviation
std <- sd(depth$depth)

# mode
mode <- Mode(depth$depth)

cutoff <- sum(mean_depth + (2*std))

# identify SNP sites with depth > mean depth + 1 standard deviation and quality score < 2x the depth at that site
temp <- bind_cols(qual, depth) %>%
  filter(depth > cutoff) %>%
  filter(qual < 2*depth) %>%
  select(locus)

write.table(temp, "data/BONNET//DEPTH.lowQloci", col.names = FALSE, row.names = FALSE, quote = FALSE)

df <- bind_cols(qual, depth) %>%
  mutate(qualcutoff = 2*depth)

removeloc <- df %>%
  filter(depth > cutoff) %>%
  filter(qual < 2*depth)

ggplot(df, aes(x = depth, y = qual)) +
  geom_point(shape = 1) +
  geom_point(data = removeloc, aes(x = depth, y = qual), shape = 1, color = "red") +
  geom_line(data = df, aes(x = depth, y = qualcutoff), color = "blue",  linetype = "dashed", size = 1) +
  geom_vline(xintercept = cutoff, color = "blue", linetype = "dashed", size = 1) +
  labs(x = "read depth (bonnetheads") +
  theme_standard

```

Mean depth per locus (across all indivuals) is `r mean_depth` and the standard deviation is `r std`. 

Filter SNP site with depth > mean depth + 1 standard deviation = `r sum(mean_depth + 2*std)` and that have quality scores < 2x the depth at that site and output depth per site.

```{bash filter depth qual II, eval=FALSE, include=FALSE}

vcftools --vcf  data/BONNET//BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQM.recode.vcf --exclude-positions data/BONNET/DEPTH.lowQloci --recode --recode-INFO-all --out data/BONNET/BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQMQUAL

vcftools --vcf data/BONNET/BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.recode.vcf --out data/BONNET/BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQMQUAL --site-mean-depth

```

Compare the distribution of mean depth per site averaged across individuals to determine cut-off value of sites with excessively high depth indicative of paralogs/multicopy loci.

```{r plot depth dist II, message=FALSE, warning=FALSE}

# calculate mean depth per site (177 individuals)
depth <- read.table("data/BONNET/BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.ldepth.mean",
           header = TRUE, stringsAsFactors = FALSE)

# mean mean depth
mean <- mean(depth$MEAN_DEPTH)

# standard deviation
std <- sd(depth$MEAN_DEPTH)

# mode
mode <- Mode(depth$MEAN_DEPTH)

cutoff <- sum(mean + (2*std))

ggplot(depth, aes(x = MEAN_DEPTH)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(MEAN_DEPTH, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mode),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = cutoff <- sum(mean + (2*std))),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per site") +
  theme_standard

```

Choose cut-off for maximum mean read depth = 120.

```{bash}

vcftools --vcf  data/BONNET/BONNET_minQ20minDP5meanDP10geno70ind50ABQDPTHMQMQUAL.recode.vcf --max-meanDP 120 --exclude-positions data/BONNET/DEPTH.lowQloci --recode --recode-INFO-all --out data/BONNET/BONNET_dDocentb

# stats BONNET
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --depth
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --site-mean-depth
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --site-quality
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --missing-indv
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --missing-site
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --het
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --freq2
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --singletons


# depth indv/locus
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --depth
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --site-mean-depth
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --missing-indv
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --indv-freq-burden
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --freq2
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --singletons
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --012

# heterozygosity per individual
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --het

# SNP call quality
vcftools --vcf data/BONNET/BONNET_dDocentb.recode.vcf --out results/BONNET_dDocentb --site-quality

```

#### Plot stats after dDocent filters

```{r stats dDocent II, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_dDocent <- read.ind.stats(dir = "results", vcf = "BONNET_dDocentb")

loc_stats_dDocent <- read.loc.stats(dir = "results", vcf = "BONNET_dDocentb")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_dDocent, aes(x = `MISS_BONNET_dDocentb`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocentb`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocentb`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocentb`, y = `MISS_BONNET_dDocentb`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocentb`, y = `MISS_BONNET_dDocentb`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_dDocent, aes(x = `Fis_BONNET_dDocentb`, y = `MEAN_DEPTH_BONNET_dDocentb`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_dDocent, aes(x = `MISS_BONNET_dDocentb`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocentb`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_dDocent, aes(x = `MEAN_DEPTH_BONNET_dDocentb`, y = `MISS_BONNET_dDocentb`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_dDocentb`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_dDocent %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_dDocent %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_dDocent) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_dDocentb`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_dDocentb.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_dDocent$`MEAN_DEPTH_BONNET_dDocentb`, site_qual$QUAL) %>%
  rename(depth = loc_stats_dDocent.MEAN_DEPTH_BONNET_dDocentb, qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m6 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

### Step 4: Filter Missing data loci/indv

Remove loci with genotype call rate < 80%.

```{bash}

vcftools --vcf data/BONNET//BONNET_dDocentb.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno80ind50 --max-missing 0.8 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno80ind50.recode.vcf --out results/BONNET_minQ20minDP5meanDP10dDocentgeno80ind50 --missing-indv

```

Identify individuals with > 30% missing data

```{r}

imiss <- read.table("results/BONNET_minQ20minDP5meanDP10dDocentgeno80ind50.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .3, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.3) %>%
  select(INDV)

write.table(LQ_indv, "data/BONNET//LQ_Ind_FIL-6a",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno80ind50.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno80ind30 --remove data/BONNET//LQ_Ind_FIL-6a --recode --recode-INFO-all

```

Remove loci with genotype call rate < 90%.

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno80ind30.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno90ind30 --max-missing 0.9 --recode --recode-INFO-all

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno90ind30.recode.vcf --out results/BONNET_minQ20minDP5meanDP10dDocentgeno90ind30 --missing-indv

```

Identify individuals with > 25% missing data

```{r}

imiss <- read.table("results/BONNET_minQ20minDP5meanDP10dDocentgeno90ind30.imiss", 
                    header = TRUE, stringsAsFactors = FALSE)

ggplot(imiss, aes(x = F_MISS)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "grey75") +
  geom_vline(xintercept = .25, color = "darkred", linetype = "dashed") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_standard

LQ_indv <- imiss %>%
  filter(F_MISS > 0.25) %>%
  select(INDV)

write.table(LQ_indv, "results/LQ_Ind_FIL-6b",
            col.names = FALSE, row.names = FALSE, quote = FALSE)

```

Remove flagged individuals

```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno90ind30.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno90ind25 --remove results/LQ_Ind_FIL-6b --recode --recode-INFO-all

```

Remove loci with genotype call rate < 95%; decompose variants and retain only SNPs.



```{bash}

vcftools --vcf data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno90ind25.recode.vcf --out data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno95ind25 --max-missing 0.95 --recode --recode-INFO-all

vcfallelicprimitives data/BONNET//BONNET_minQ20minDP5meanDP10dDocentgeno95ind25.recode.vcf --keep-info --keep-geno > data/BONNET//SNPs.vcf

vcftools --vcf data/BONNET//SNPs.vcf --out data/BONNET//BONNET_FIL-6 --remove-indels --recode --recode-INFO-all


# depth indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --depth
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --site-mean-depth
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --geno-depth

# missing data indv/locus
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --missing-indv
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --missing-site

# allele freq/indv freq buden
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --indv-freq-burden
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --freq2
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --singletons
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --012

# heterozygosity per individual
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --het

# SNP call quality
vcftools --vcf data/BONNET//BONNET_FIL-6.recode.vcf --out results/BONNET_FIL-6 --site-quality

```

### Plot stats for filtered data set

```{r stats FIL 6, fig.height=20, fig.width=10, message=FALSE, warning=FALSE}

# load stats files ----
ind_stats_FIL_6 <- read.ind.stats(dir = "results", vcf = "BONNET_FIL-6")

loc_stats_FIL_6 <- read.loc.stats(dir = "results", vcf = "BONNET_FIL-6")

# plot missing data per indv ----
p1 <- ggplot(ind_stats_FIL_6, aes(x = `MISS_BONNET_FIL-6`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "missing data per indv") +
  theme_standard

# plot Fis per indv ----
p2 <- ggplot(ind_stats_FIL_6, aes(x = `Fis_BONNET_FIL-6`)) +
  geom_histogram(binwidth = .01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv") +
  theme_standard

# plot read depth per indv ----
p3 <- ggplot(ind_stats_FIL_6, aes(x = `MEAN_DEPTH_BONNET_FIL-6`)) +
  geom_histogram(binwidth = 10, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per indv") +
  theme_standard

# plot depth vs missing ----
p4 <- ggplot(ind_stats_FIL_6, aes(x = `MEAN_DEPTH_BONNET_FIL-6`, y = `MISS_BONNET_FIL-6`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per indv", y = "% missing data") +
  theme_standard

# plot Fis vs missing data per indv ----
p5 <- ggplot(ind_stats_FIL_6, aes(x = `Fis_BONNET_FIL-6`, y = `MISS_BONNET_FIL-6`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.5),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "% missing data") +
  theme_standard

# plot Fis vs mean depth per indv ----
p6 <- ggplot(ind_stats_FIL_6, aes(x = `Fis_BONNET_FIL-6`, y = `MEAN_DEPTH_BONNET_FIL-6`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`Fis_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MEAN_DEPTH_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "Fis per indv", y = "mean depth per indv") +
  theme_standard

# plot distribution missing data per locus ----
p7 <- ggplot(loc_stats_FIL_6, aes(x = `MISS_BONNET_FIL-6`)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MISS_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "% missing data per locus") +
  theme_standard

# plot distribution mean read depth ----
p8 <- ggplot(loc_stats_FIL_6, aes(x = `MEAN_DEPTH_BONNET_FIL-6`)) +
  geom_histogram(binwidth = 5, color = "black", fill = "grey95") +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean read depth per locus") +
  theme_standard

# plot read depth vs missing data ----
p9 <- ggplot(loc_stats_FIL_6, aes(x = `MEAN_DEPTH_BONNET_FIL-6`, y = `MISS_BONNET_FIL-6`)) +
  geom_point() +
  geom_vline(aes(xintercept = mean(`MEAN_DEPTH_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(`MISS_BONNET_FIL-6`, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 0.1),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "% missing data") +
  theme_standard

# plot no of SNPs per locus ----
p10 <- loc_stats_FIL_6 %>%
  count(CHR) %>%
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey95") + 
  labs(x = "number of SNPs per locus") +
  theme_standard

temp <- loc_stats_FIL_6 %>%
  count(CHR)

# plot number of SNPs per contig vs. mean depth ----
p11 <- left_join(temp, loc_stats_FIL_6) %>%
  ggplot() +
  geom_point(aes(x = n, y = `MEAN_DEPTH_BONNET_FIL-6`)) +
  labs(x = "number of SNPs per contig", y = "mean depth") +
  theme_standard

# plot depth vs SNP quality ----
site_qual <- read.table("results/BONNET_FIL-6.lqual", 
                        header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(PROB = 10^(-QUAL/10))

temp <- data.frame(loc_stats_FIL_6$`MEAN_DEPTH_BONNET_FIL-6`, site_qual$QUAL) %>%
  rename(depth = loc_stats_FIL_6..MEAN_DEPTH_BONNET_FIL.6., qual = site_qual.QUAL)

p12 <- ggplot(temp, aes(x = depth, y = qual)) +
  geom_point(size = 1) +
  geom_vline(aes(xintercept = mean(depth, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = mean(qual, na.rm = TRUE)),
                 color = "red", linetype = "dashed", size = 1) +
  geom_hline(aes(yintercept = 20),
                 color = "darkblue", linetype = "dashed", size = 1) +
  labs(x = "mean depth per locus", y = "SNP quality") +
  theme_standard

m8 <- multiplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, cols=2)

```

Data set contains `r nrow(ind_stats_FIL_6)` individuals and `r nrow(loc_stats_FIL_6)` loci.

## Run SNP counting script

```{bash, eval=FALSE, include=FALSE}

# run in BONNET folder

echo "FILTER SNP CONTIG INDV" > Filter.count

for i in *.vcf
do
  SNP=$(grep -cv '#' $i)
  CONTIG=$(grep -v '#' $i | cut -f 1 | sort | uniq | wc -l)
  INDV=$(vcfsamplenames $i | wc -l)
  echo "$i $SNP $CONTIG $INDV" >> Filter.count
done

```

Compare SNPs/contigs/indv at each filtering step and between filtering schemes.

```{r}

count <- read.table("data/BONNET/Filter.count", 
                    header = TRUE, stringsAsFactors = FALSE)
